---
title: "MethylFlash"
author: "Maggie Schedl"
date: "4/6/2021"
output: github_document
---
Load in Libraries
```{r}
library(ggplot2)
library(ggpmisc)
```
Check directory
```{r}
getwd()
```

# For generating the Standard Curve I went through a couple of options
# I tried 3 types of regression lines.  There is a standard linear regression, a quadratic regression ("Use this method when the standard curve is flat due to a saturated signal intensity at high PC%"), and a standard linear regression with the last points (5% PC) removed because the manual says that not all points are necessary: "determine the slope (OD/%) of the standard curve using linear regression and the most linear part (at least 4 concentration points incluing 0) of the standard curve for optimal slope calculation" 
# All of the R^2 and significances for the regressions are high/significant so it probably doesn't matter that much which way we chose to analyze it. 
# There is no difference in the slopes of any of the lines if I used the averages of the duplicates, and the slopes are what determine the calculations 



# Read in duplicate absorbances of the standards 
```{r}
#read in csv of absorbances
Standards <- read.csv("PC_Standard_Curve.csv")
Standards <- Standards[-c(15), ] # remove row with NA, not sure why it inputs like that
```

Read in csv of the average OD for each sample
```{r}
MethylFlashAvgOD <- read.csv("MethylFlash-20210402-Average.csv")
```

Make basic plot of the standards to see what it looks like
```{r}
ggplot(Standards, aes(x=Percent_PC, y=Absorbance_450)) + geom_point()
```
You can see some sort of plateau here at the last points, although there is a lot of variability there. 

# Linear Regression for Duplicate Standards
```{r}
ggplot(Standards, aes(x=Percent_PC, y=Absorbance_450)) + geom_point() + geom_smooth(method=lm, se=FALSE) + stat_poly_eq(formula = y ~ x , aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = TRUE) 

# what is the model?
model_lm <- lm(Absorbance_450 ~ Percent_PC, data = Standards)
summary(model_lm) # look at statistics
```
Using the R^2 and p value, this is a good fit. 

Calculation of %5-mC using the linear regression model for duplicate standards

```{r}
# Linear Model Calculation
# Equation is 5-mc% = ((Average Sample OD - Average NC OD)/ (Slope * 100ng)) * 100
AverageNC <- 0.0445 # average OD of negative control 
SlopeLM <- 0.22044 # from linear model

MethylFlashAvgOD$LM_5mc_Percentage <- (((MethylFlashAvgOD$Average_OD-AverageNC)/(SlopeLM * 100))*100)
```


Second Order Regression or Quadratic Model because the high percentage PC flatten

```{r}
# try fitting quadratic model
#create squared variable
Standards$Percent_PC_2 <- Standards$Percent_PC^2
#do model
quadraticModel <- lm(Absorbance_450 ~ Percent_PC + Percent_PC_2, data=Standards)
summary(quadraticModel) # look at statistics

# plot
ggplot(Standards, aes(x=Percent_PC, y=Absorbance_450)) + geom_point() + stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, se=FALSE) + stat_poly_eq(formula = y ~ x + I(x^2), aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = TRUE)
```

Calculation of 5mC% using the quadratic model with duplicate standards 
```{r}
# Equation for quadratic is: Y = aX^2 + bX 
# a and b could also be called Slope1 and Slope2

# Equation is 5-mC% = ((((Slope2^2 + 4Slope1(Average Sample OD - Average NC OD)^0.5) - Slope2 )/2Slope1)/100)*100

Slope1 <- -0.037307
Slope2 <- 0.4067

MethylFlashAvgOD$QU_5mc_Percentage <- (((((Slope2^2+(4*Slope1*(MethylFlashAvgOD$Average_OD-AverageNC)))^.5)-Slope2)/(2*Slope1))/100)*100
```

Linear Model without 5% PC Standard Point

```{r}
# Remove 5% points

Standards_Modified <- Standards[-c(13,14), ] #remove last two rows

ggplot(Standards_Modified, aes(x=Percent_PC, y=Absorbance_450)) + geom_point() + geom_smooth(method=lm, se=FALSE) + stat_poly_eq(formula = y ~ x , aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = TRUE) 
model_lm2 <- lm(Absorbance_450 ~ Percent_PC, data = Standards_Modified)
summary(model_lm2)  
```

Calculate 5mC% using the modified linear model using duplicate standards
```{r}
# Equation is 5-mc% = ((Average Sample OD - Average NC OD)/ (Slope * 100ng)) * 100
SlopeLM2 <- 0.33731

MethylFlashAvgOD$LM_2__5mc_Percentage <- (((MethylFlashAvgOD$Average_OD-AverageNC)/(SlopeLM2 * 100))*100)

```


Print the 5mC% calculations

```{r}
print(MethylFlashAvgOD)
```













